Analyzing X-ray images in Python with scikit-image

### 2016-08-21 skimage_tomo.tex
Further minor edits

Several software applications and libraries are available to synchrotron
users to process their images. ImageJ \citep{Abramoff2004, Schneider2012}
and its distribution Fiji \citep{Schindelin2012} is a popular tool for
2-D and 3-D images, thanks to its intuitive menus and graphical tools,
and the wealth of plugins contributed by a vivid
community \citep{Schindelin2015}. For 3-D images, commercial tools such
as Avizo 3D software (TM) are appreciated for an intuitive graphical
pipeline and advanced 3D visualization. Alternatively, the use of a
programming language gives finer control and more possibilities, provided
classical processing algorithms can be called from libraries -- hence
limiting the programming task and the risk of bugs. Matlab (TM) and its
image processing toolbox are popular in the academic community of
computer vision and image processing. The Python language is widely used
in the scientific world and in synchrotron facilities. As a
general-purpose language, Python is used in synchrotrons to control
device servers \citep{pytango}, to access raw data of X-ray
detectors \citep{Knudsen2013}, to reconstruct tomography volumes from
radiographs \citep{Gursoy2014, Mirone2014}, and in data
processing packages for macromolecular cristallography \citep{Adams2010},
azimuthal integration of diffraction data \citep{Ashiotis2015} or
fluorescence analysis \citep{pymca}.

@@

\begin{figure*}
    \centerline{\includegraphics[width=0.8\textwidth]{ecosystem_landscape}}
    \caption{\csentence{\texttt{scikit-image} and the Scientific Python
	ecosystem}. Images are opened from files as \texttt{NumPy}
	arrays. Functions of \texttt{scikit-image} transform image arrays
	into other arrays with the same dimensions, or into arrays of
	numbers corresponding to features of the image. The output of
	scikit-image functions can be passed to other Python modules
	relying on \texttt{NumPy} arrays, such as \texttt{Scipy} or
	\texttt{scikit-learn}. Image-shaped arrays are transformed into
	visualizations with \texttt{matplotlib} (2D) or \texttt{Mayavi}
	(3D). A variety of environments is available for code development
	and execution, from classical IDEs to Jupyter notebooks.
    \label{fig:ecosystem}}
\end{figure*}

@@

Most functions take indifferently 2-D or 3-D images as arguments, so that
the same functions can be used to process tomography and microscopy
images. Only a few functions do not support 3-D images and raise an error
when passed a 3-D argument:
\begin{lstlisting}
>>> filters.prewitt(im)
[...]
ValueError: The parameter `image` must be a 2-dimensional array
\end{lstlisting}

Another type of functions returns numerical value(s) that do not
correspond to image objects, such as pixel coordinates of objects of
interest or statistical
information about the image:
\begin{lstlisting}
>>> from skimage import exposure
>>> counts, bins = exposure.histogram(im)
>>> counts.shape
(256,)
\end{lstlisting}

@@

While several interpreters are available to execute interactively Python
instructions and scripts, the most popular one in the scientific world is
the IPython interpreter \citep{Perez2007, Rossant2015}. IPython is an
advanced interpreter, which integrates syntax highlighting, text auto-completion,
a debugger,
introspection and profiling methods, and online help. Several
Integrated Development Environments (IDEs) include IPython,
together with other components such as a text editor. Notable examples
include Spyder (Fig.~\ref{fig:spyder}), PyCharm, and Visual Studio Code.

@@

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
from the IPython project. Jupyter notebooks provide an interactive development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (FIGURE).
(jni: FURTHER TEXT ABOUT IMPACT OF JUPYTER NOTEBOOKS)

\paragraph{NumPy arrays} are the cornerstone of the Scientific Python
ecosystem, and of \texttt{scikit-image} operations in particular.
Cropping or downsampling an image, or retrieving pixels corresponding to
a given label in a segmentation are all one-liners NumPy code. To
illustrate the compactness of NumPy code, modifying pixels values below a
threshold can be written as
\begin{lstlisting}
im[im < 0.5] = 0
\end{lstlisting}
using the possibility to index arrays with boolean arrays, called
\emph{masking}. Also, \texttt{NumPy} uses memory sparingly and avoids
making new copies of arrays whenever possible, which is a important
requirement when dealing with the gigabyte-sized images of tomography.
For example, cropping a subvolume as follows does not create a copy of the
original array:
\begin{lstlisting}
sub_volume = im[100:-100, 100:-100, 100:-100]
\end{lstlisting}

>>>>

Several software applications and libraries are available to synchrotron
users to process their images. ImageJ \citep{Abramoff2004, Schneider2012}
and its distribution Fiji \citep{Schindelin2012} is a popular tool for
2-D and 3-D images, thanks to its intuitive menus and graphical tools,
and the wealth of plugins contributed by a vivid
community \citep{Schindelin2015}. For 3-D images, commercial tools such
as Avizo 3D software (TM) are appreciated for an intuitive graphical
pipeline and advanced 3D visualization. Alternatively, the use of a
programming language gives finer control and more possibilities, provided
classical processing algorithms can be called from libraries -- hence
limiting the programming task and the risk of bugs. Matlab (TM) and its
image processing toolbox are popular in the academic community of
computer vision and image processing. The Python language is widely used
in the scientific world and in synchrotron facilities. As a
general-purpose language, Python is used in synchrotrons to control
device servers \citep{pytango}, to access raw data of X-ray
detectors \citep{Knudsen2013}, to reconstruct tomography volumes from
radiographs \citep{Gursoy2014, Mirone2014}, and in data
processing packages for macromolecular cristallography \citep{Adams2010},
azimuthal integration of diffraction data \citep{Ashiotis2015}, or
fluorescence analysis \citep{pymca}.

@@

\begin{figure*}
    \centerline{\includegraphics[width=0.8\textwidth]{ecosystem_landscape}}
    \caption{\csentence{\texttt{scikit-image} and the Scientific Python
	ecosystem}. Images are opened from files as \texttt{NumPy}
	arrays. Functions of \texttt{scikit-image} transform image arrays
	into other arrays with the same dimensions, or into arrays of
	numbers corresponding to features of the image. The output of
	scikit-image functions can be passed to other Python modules
	relying on \texttt{NumPy} arrays, such as \texttt{SciPy} or
	\texttt{scikit-learn}. Image-shaped arrays are transformed into
	visualizations with \texttt{matplotlib} (2D) or \texttt{Mayavi}
	(3D). A variety of environments is available for code development
	and execution, from classical IDEs to Jupyter notebooks.
    \label{fig:ecosystem}}
\end{figure*}

@@

Most functions transparently handle 2-D, 3-D, or even higher-dimensional
images as arguments, so
the same functions can be used to process tomography, microscopy, or natural
images. A few functions do not support 3-D images and raise an error
when passed a 3-D argument:
\begin{lstlisting}
>>> filters.prewitt(im)
[...]
ValueError: The parameter `image` must be a 2-dimensional array
\end{lstlisting}

However, the proportion of functions supporting 3D images is always increasing,
thanks to the many contributors to the library.

A subset of functions returns not images but other value types,
such as pixel coordinates of objects of interest, or statistical
information about the image:
\begin{lstlisting}
>>> from skimage import exposure
>>> counts, bins = exposure.histogram(im)
>>> counts.shape
(256,)
\end{lstlisting}

@@

While several interpreters are available to execute Python
instructions and scripts, the most popular one in the scientific world is
the IPython interpreter \citep{Perez2007, Rossant2015}. IPython is an
advanced interpreter, which integrates syntax highlighting, text auto-completion,
a debugger,
introspection and profiling methods, and online help. Several
Integrated Development Environments (IDEs) come bundled with IPython,
together with other components such as a text editor. Notable examples
include Spyder (Fig.~\ref{fig:spyder}), PyCharm, and Visual Studio Code.

@@

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
from the IPython project. Jupyter notebooks provide a development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (FIGURE).
As of July 2016, more than 500,000 Jupyter notebooks were posted on GitHub,
demonstrating their wide adoption by the community as workflow-sharing
tools (http://archive.ipython.org/media/SciPy2016JupyterLab.pdf).

\paragraph{NumPy arrays} are the cornerstone of the Scientific Python
ecosystem, and of \texttt{scikit-image} operations in particular.
Cropping or downsampling an image, or retrieving pixels corresponding to
a given label in a segmentation are all one-liners NumPy code. To
illustrate the compactness of NumPy code, modifying pixels values below a
threshold can be written as
\begin{lstlisting}
im[im < 0.5] = 0
\end{lstlisting}
using the ability to index arrays with boolean arrays, called
\emph{masking}. \texttt{NumPy} uses memory sparingly and avoids
making new copies of arrays whenever possible, an important
requirement when dealing with the gigabyte-sized images of tomography.
For example, cropping a subvolume as follows does not create a copy of the
original array:
\begin{lstlisting}
sub_volume = im[100:-100, 100:-100, 100:-100]
\end{lstlisting}
